### 2020-11-30 paper/localpsf.tex
edited abstrach

% REQUIRED
\begin{abstract}
  We present an efficient matrix-free method for approximating smoothly varying blur operators, and apply the method to approximate Hessians in large-scale inverse problems with highly informative data. Such Hessians are high rank and therefore expensive to approximate with existing methods. We compute the impulse responses of the operator at many points by applying the operator to a small number of Dirac combs of point sources. Within each Dirac comb, the point sources must be spaced sufficiently far apart so that the impulse responses to those point sources do not overlap with each other or with the boundary. We estimate the required spacing of point sources by applying the operator to a small number of constant, linear, and quadratic functions, then postprocessing the results. Once computed, the impulse responses are used to form a low-rank approximation of the spatially varying impulse response function (SVIR), using a continuous analog of the interpolative matrix decomposition. Smoothly varying blur operators have a low rank SVIR, even if the operator itself is high rank. We convert the resulting approximation to hierarchical matrix format, where it is inverted using fast hierarchical matrix arithmetic and used to solve linear systems.  We apply the method to a large-scale PDE-constrained inverse problem with highly informative data. 
%  {\color{red}Numerical results demonstrate that our method substantially outperforms existing methods.}
\end{abstract}

@@

Let $\Omega \subset \mathbb{R}^d$ be a bounded domain, let $a:L^2(\Omega) \times L^2(\Omega) \rightarrow \mathbb{R}$,
\begin{equation}
	\label{eq:kernel_representation}
	a(v,u) := \int_\Omega \int_\Omega v(y) A(y,x) u(x) dx dy,
\end{equation}
be a bilinear form with integral kernel $A:\Omega \times \Omega \rightarrow \mathbb{R}$, and let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be the corresponding linear operator generated by currying. That is, for each $u\in L^2(\Omega)$, the linear functional $\mathcal{A}u:L^2(\Omega)\rightarrow \mathbb{R}$ is defined by
\begin{equation*}
	(\mathcal{A}u)(v) := a(v,u).
\end{equation*}
We seek to approximate operators $\mathcal{A}$ in the ``matrix free'' setting, in which access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$. Here ($\mathcal{A}^Tv)(u) := (\mathcal{A}u)(v)$. Integral kernel representation \eqref{eq:kernel_representation}, is assumed to exist,
%e.g., by the Schwarz kernel theorem,
but one cannot easily access entries $A(y,x)$ of the kernel. Such operators arise, for example, as Schur complements in reduced space methods for solving partial differential equations (PDEs), and as Hessians in distributed parameter PDE-constrained inverse problems.

>>>>

% REQUIRED
\begin{abstract}
  We present an efficient matrix-free method for approximating locally translation invariant operators with non-negative integral kernels. The idea of the method is to compute impulse responses of the operator at a collection of well-spaced points, then interpolate these impulse responses to approximate impulse responses at other arbitrary points. We convert the resulting approximation to hierarchical matrix format, which allows us to perform matrix inversion or other matrix operations using fast hierarchical matrix arithmetic. Impulse responses are computed by applying the operator to a small number of Dirac comb ``batches'' of point sources, and special care is taken to ensure that the supports of the impulse responses within each batch do not overlap. We apply the method to approximate Hessians in large-scale PDE-constrained inverse problem with highly informative data. Numerical results demonstrate that our method substantially outperforms existing Hessian approximation methods which are based on low-rank approximation. Our method is able to form high quality approximations of high rank Hessians using only a small number of Hessian matrix-vector products.
%  Within each Dirac comb, the point sources must be spaced sufficiently far apart so that the impulse responses to those point sources do not overlap. We estimate the required spacing of point sources by applying the operator to a small number of constant, linear, and quadratic functions, then postprocessing the results.
\end{abstract}

@@

Let $\Omega \subset \mathbb{R}^d$ be a bounded domain, and let $a:L^2(\Omega) \times L^2(\Omega) \rightarrow \mathbb{R}$,
\begin{equation}
	\label{eq:kernel_representation}
	a(v,u) := \int_\Omega \int_\Omega v(y) A(y,x) u(x) dx dy,
\end{equation}
be a bilinear form with non-negative integral kernel $A:\Omega \times \Omega \rightarrow \mathbb{R}$, $A(y,x) \ge 0$. Let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be the corresponding linear operator generated by currying. That is, for each $u\in L^2(\Omega)$, the linear functional $\mathcal{A}u:L^2(\Omega)\rightarrow \mathbb{R}$ is defined by
\begin{equation*}
	(\mathcal{A}u)(v) := a(v,u).
\end{equation*}
We seek to approximate operators $\mathcal{A}$ in the ``matrix free'' setting, in which access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$. Here ($\mathcal{A}^Tv)(u) := (\mathcal{A}u)(v)$. Integral kernel representation \eqref{eq:kernel_representation}, is assumed to exist,
%e.g., by the Schwarz kernel theorem,
but one cannot easily access entries $A(y,x)$ of the kernel. Such operators arise, for example, as Schur complements in reduced space methods for solving partial differential equations (PDEs), and as Hessians in distributed parameter PDE-constrained inverse problems.
